---
title: "Climate"
author: "Elise Boos, Kelsey Husted, Taro Katayama"
date: "2023-04-11"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

###Regional Weather Forecasting

##Objectives

Forecasting temperatures is useful for energy usage projections in the future. Additionally, analyzing climate-related data at a national-level is important since climate change may impact certain regions more significantly than others. This could potentially result in energy allotment adjustments.

##Method

Temperature data was obtained from the International Energy Agency website and was available to download at the grid, national, and sub-national level. Daily and monthly resolution data was provided that ranged from 2000 to the present. After importing and wrangling the data below, a time series variable was created.

(Discussion and Limitations at the bottom of rmd)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load Data

```{r, warning=FALSE, message=FALSE}
require(tidyverse)
require(tseries)
require(forecast)
require(lubridate)
require(ggplot2)
require(Kendall)
library(kableExtra)
temp <- read.csv("./Data_Raw/Temperature_Data.csv")
```


###Tidy data

```{r  warning=FALSE, message=FALSE}
temp_processed<- temp%>%
  mutate(Date = as.Date(Date, format = "%d-%m-%y"))%>%
  rename(Temp= Temperaturedaily)%>%
  arrange(Date, Country)%>%
  select(Date, Country, Temp)

```


###Time series transformation (BRA)

```{r  warning=FALSE, message=FALSE}
#Filter for Brazil 
#Temperature
BRA_temp <- temp_processed%>%
  filter(Country == 'BRA')


fday <- first(BRA_temp$Date)

#Time series transformation ts()
ts_temp_BRA <- ts(BRA_temp[,3], frequency = 365, start = c(year(fday), month(fday), day(fday)))

#Time series transformation msts()
ts_temp_BRA2 <- msts(BRA_temp[,3], seasonal.periods =c(7,365.25),
                           start=c(year(fday), month(fday), day(fday)))

#Decompose
decompose_temp_BRA <- decompose(ts_temp_BRA, type = "additive")
plot(decompose_temp_BRA)

ggplot(BRA_temp, aes(x=Date, y =Temp))+
  geom_line()+
  geom_smooth(method = lm)

summary(SeasonalMannKendall(ts_temp_BRA))
```


###Time series transformation (USA)

```{r  warning=FALSE, message=FALSE}
#Filter for USA
#Temperature
USA_temp <- temp_processed%>%
  filter(Country == 'USA')


#Time series transformation
ts_temp_USA <- ts(USA_temp[,3], frequency = 365, start = c(2000,1))

#Time series transformation msts()
ts_temp_USA2 <- msts(USA_temp[,3], seasonal.periods =c(7,365.25),
                           start=c(year(fday), month(fday), day(fday)))
#Decompose
decompose_temp_USA <- decompose(ts_temp_USA, type = "additive")
plot(decompose_temp_USA)

ggplot(USA_temp, aes(x=Date, y =Temp))+
  geom_line()+
  geom_smooth(method = lm)

summary(SeasonalMannKendall(ts_temp_USA))
```


###Time series transformation (AUS)

```{r  warning=FALSE, message=FALSE}
#Filter for Australia
#Temperature
AUS_temp <- temp_processed%>%
  filter(Country == 'AUS')

#Time series transformation
ts_temp_AUS <- ts(AUS_temp[,3], frequency = 365, start = c(2000,1))

#Time series transformation msts()
ts_temp_AUS2 <- msts(AUS_temp[,3], seasonal.periods =c(7,365.25),
                           start=c(year(fday), month(fday), day(fday)))

#Decompose
decompose_temp_AUS <- decompose(ts_temp_AUS, type = "additive")
plot(decompose_temp_AUS)

ggplot(AUS_temp, aes(x=Date, y =Temp))+
  geom_line()+
  geom_smooth(method = lm)

summary(SeasonalMannKendall(ts_temp_AUS))
```


###Temperature Forecasting Test (BRA)

```{r forecast test  warning=FALSE, message=FALSE}
#AUTO ARIMA
ts_temp_BRA_sub <-window(ts_temp_BRA, start = c(2000,1), end = c(2021,12))

arima <- auto.arima(ts_temp_BRA_sub,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast <- forecast(arima,h=365)

#ARIMA with seasonality added back

seasonality<- decompose_temp_BRA$seasonal[1:365]

for_and_seasonality<- ARIMA_forecast$mean + seasonality

ts_temp_BRA2_sub <-window(ts_temp_BRA2, start = c(2000,1), end = c(2021,12))
### ARIMA + Fourier Terms
ARIMA_Four <- auto.arima(ts_temp_BRA2_sub,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_BRA2_sub,
                                      K=c(2,12)))
ARIMA_Four_for <- forecast(ARIMA_Four,
                           xreg=fourier(ts_temp_BRA2_sub,
                                        K=c(2,12),
                                        h=365),
                           h=365)


#### NEURAL NETWORK 
NN_fit <- nnetar(ts_temp_BRA2_sub,p=1,P=1)

#5 year forecast
NN_for <- forecast(NN_fit, h=365)


### TBATS
TBATS_fit <- tbats(ts_temp_BRA2_sub)

TBATS_for <- forecast(TBATS_fit, h=365)

### STL + ETS
ETS_fit <- stlf(ts_temp_BRA2_sub, h=365)

###ACCURACY
a_score <- accuracy(ARIMA_forecast)
a2_score <- accuracy(ARIMA_Four_for)
NN_score <- accuracy(NN_for)
TBATS_score <- accuracy(TBATS_for)
ETS_score <- accuracy(ETS_fit)

scores <- as.data.frame(rbind(a_score, a2_score, NN_score, TBATS_score, ETS_score))
row.names(scores) <- c("ARIMA", "ARIMA FOURIER","NN", "TBATS", "STL + ETS")
scores


autoplot(window(ts_temp_BRA, start = c(2019,1))) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  autolayer(TBATS_for,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)+
  autolayer(ARIMA_forecast, series="ARIMA",PI=FALSE)+
  autolayer(for_and_seasonality, series = "ARIMA + Seasonality")+
  ylab("Temperature") 
```


###Temperature Forecasting (BRA)

```{r forecast  warning=FALSE, message=FALSE}
#AUTO ARTIMA
arima <- auto.arima(ts_temp_BRA,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast <- forecast(arima,h=365)

#ARIMA with seasonality added back

seasonality<- decompose_temp_BRA$seasonal[1:365]

for_and_seasonality<- ARIMA_forecast$mean + seasonality

### ARIMA + Fourier Terms
ARIMA_Four <- auto.arima(ts_temp_BRA2,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_BRA2,
                                      K=c(2,12)))
ARIMA_Four_for <- forecast(ARIMA_Four,
                           xreg=fourier(ts_temp_BRA2,
                                        K=c(2,12),
                                        h=365),
                           h=365)

#### NEURAL NETWORK 
NN_fit <- nnetar(ts_temp_BRA2,p=1,P=1)

#5 year forecast
NN_for <- forecast(NN_fit, h=365)


### TBATS
TBATS_fit <- tbats(ts_temp_BRA2)

TBATS_for <- forecast(TBATS_fit, h=365)

### STL + ETS
ETS_fit <- stlf(ts_temp_BRA2, h=365)


###PLOT
autoplot(window(ts_temp_BRA, start = c(2019,1))) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  autolayer(ARIMA_forecast, series="ARIMA",PI=FALSE)+
  autolayer(TBATS_for,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)+
  autolayer(for_and_seasonality, series="ARIMA + Seasonality")+
  ylab("Temperature") 

#create accuracy test for each forecast
a_score <- accuracy(ARIMA_forecast)
a2_score <- accuracy(ARIMA_Four_for)
NN_score <- accuracy(NN_for)
TBATS_score <- accuracy(TBATS_for)
ETS_score <- accuracy(ETS_fit)


#create accuracy test matrix
scores <- as.data.frame(rbind(a_score, a2_score, NN_score, TBATS_score, ETS_score))
row.names(scores) <- c("ARIMA", "ARIMA FOURIER","NN", "TBATS", "STL + ETS")
scores

kbl(scores, 
      caption = "Forecast Accuracy for Brazil Temperature Data",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", html_font="Cambria")
```


###Temperature Forecasting Test (AUS)

```{r forecast test AUS  warning=FALSE, message=FALSE}
#AUTO ARTIMA
ts_temp_AUS_sub <-window(ts_temp_AUS, start = c(2000,1), end = c(2021,12))

arima_AUS <- auto.arima(ts_temp_AUS_sub,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast_AUS <- forecast(arima_AUS,h=365)


ts_temp_AUS2_sub <-window(ts_temp_AUS2, start = c(2000,1), end = c(2021,12))
### ARIMA + Fourier Terms
ARIMA_Four_AUS <- auto.arima(ts_temp_AUS2_sub,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_AUS2_sub,
                                      K=c(2,12)))
ARIMA_Four_AUS_for <- forecast(ARIMA_Four_AUS,
                           xreg=fourier(ts_temp_AUS2_sub,
                                        K=c(2,12),
                                        h=365),
                           h=365)

#### NEURAL NETWORK 
NN_fit_AUS <- nnetar(ts_temp_AUS2_sub,p=1,P=1)

#5 year forecast
NN_for_AUS <- forecast(NN_fit_AUS, h=365)


### TBATS
TBATS_fit_AUS <- tbats(ts_temp_AUS2_sub)

TBATS_for_AUS <- forecast(TBATS_fit_AUS, h=365)

### STL + ETS
ETS_fit_AUS <- stlf(ts_temp_AUS2_sub, h=365)

###ACCURACY
a_score_AUS <- accuracy(ARIMA_forecast_AUS)
a2_score_AUS <- accuracy(ARIMA_Four_AUS_for)
NN_score_AUS <- accuracy(NN_for_AUS)
TBATS_score_AUS <- accuracy(TBATS_for_AUS)
ETS_score_AUS <- accuracy(ETS_fit_AUS)


scores_AUS <- as.data.frame(rbind(a_score_AUS, a2_score_AUS, NN_score_AUS, TBATS_score_AUS, ETS_score_AUS))
row.names(scores_AUS) <- c("ARIMA", "ARIMA FOURIER","NN", "TBATS", "STL + ETS")
scores_AUS


autoplot(window(ts_temp_AUS, start = c(2019,1))) +
  autolayer(NN_for_AUS, series="Neural Network",PI=FALSE)+
  autolayer(TBATS_for_AUS,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit_AUS,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_AUS_for, series="ARIMA_FOURIER",PI=FALSE)+
  autolayer(ARIMA_forecast_AUS, series="ARIMA",PI=FALSE)+
  ylab("Temperature") 
```


###Temperature Forecasting (AUS)

```{r forecast AUS  warning=FALSE, message=FALSE}
#AUTO ARTIMA
arima_AUS <- auto.arima(ts_temp_AUS,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast <- forecast(arima_AUS,h=365)

### ARIMA + Fourier Terms
ARIMA_Four <- auto.arima(ts_temp_AUS2,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_AUS2,
                                      K=c(2,12)))
ARIMA_Four_for <- forecast(ARIMA_Four,
                           xreg=fourier(ts_temp_AUS2,
                                        K=c(2,12),
                                        h=365),
                           h=365)

#### NEURAL NETWORK 
NN_fit <- nnetar(ts_temp_AUS2,p=1,P=1)

#5 year forecast
NN_for <- forecast(NN_fit, h=365)


### TBATS
TBATS_fit <- tbats(ts_temp_AUS2)

TBATS_for <- forecast(TBATS_fit, h=365)

### STL + ETS
ETS_fit <- stlf(ts_temp_AUS2, h=365)


###PLOT
autoplot(window(ts_temp_AUS, start = c(2019,1))) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  autolayer(ARIMA_forecast, series="ARIMA",PI=FALSE)+
  autolayer(TBATS_for,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)+
  ylab("Temperature") 


a_score <- accuracy(ARIMA_forecast)
a2_score <- accuracy(ARIMA_Four_for)
NN_score <- accuracy(NN_for)
TBATS_score <- accuracy(TBATS_for)
ETS_score <- accuracy(ETS_fit)


scores <- as.data.frame(rbind(a_score, a2_score, NN_score, TBATS_score, ETS_score))
row.names(scores) <- c("ARIMA", "ARIMA FOURIER","NN", "TBATS", "STL + ETS")
scores

kbl(scores, 
      caption = "Forecast Accuracy for Australia Temperature Data",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", html_font="Cambria")
```

###Temperature Forecasting Test (USA)

```{r forecast test USA}
#AUTO ARTIMA
ts_temp_sub <-window(ts_temp_USA, start = c(2000,1), end = c(2020,12))

ts_actual_forecast<- window(ts_temp_USA, start = c(2021,1), end = c(2021,12))

arima <- auto.arima(ts_temp_sub,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast <- forecast(arima,h=365)


ts_temp_USA2_sub <-window(ts_temp_USA2, start = c(2000,1), end = c(2020,12))
### ARIMA + Fourier Terms
ARIMA_Four <- auto.arima(ts_temp_USA2_sub,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_USA2_sub,
                                      K=c(2,12)))
ARIMA_Four_USA_for <- forecast(ARIMA_Four,
                           xreg=fourier(ts_temp_USA2_sub,
                                        K=c(2,12),
                                        h=365),
                           h=365)

#### NEURAL NETWORK 
NN_fit <- nnetar(ts_temp_USA2_sub,p=1,P=1)

#5 year forecast
NN_for <- forecast(NN_fit, h=365)


### TBATS
TBATS_fit <- tbats(ts_temp_USA2_sub)

TBATS_for <- forecast(TBATS_fit, h=365)

### STL + ETS
ETS_fit <- stlf(ts_temp_USA2_sub, h=365)



###ACCURACY
a_score <- accuracy(ARIMA_forecast, ts_actual_forecast)
a2_score <- accuracy(ARIMA_Four_USA_for, ts_actual_forecast)
NN_score <- accuracy(NN_for, ts_actual_forecast)
TBATS_score <- accuracy(TBATS_for, ts_actual_forecast)
ETS_score <- accuracy(ETS_fit, ts_actual_forecast)


scores <- as.data.frame(rbind(a_score, a2_score, NN_score, TBATS_score, ETS_score))
row.names(scores) <- c("ARIMA_training", "ARIMA_test", "ARIMA FOURIER_training","ARIMA FOURIER_TEST",
"NN_training", "NN_test","TBATS_training", "TBATS_test", "STL + ETS_training", "STL + ETS_test")
scores


autoplot(window(ts_temp_USA, start = c(2019,1))) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  autolayer(TBATS_for,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)+
  autolayer(ARIMA_forecast, series="ARIMA",PI=FALSE)+
  ylab("Temperature") 
```


##Temperature Forecasting (USA)

```{r forecast USA  warning=FALSE, message=FALSE}
#AUTO ARTIMA
arima <- auto.arima(ts_temp_USA,seasonal=FALSE, 
                             lambda=0)

ARIMA_forecast <- forecast(arima,h=365)

### ARIMA + Fourier Terms
ARIMA_Four <- auto.arima(ts_temp_USA2,
                         seasonal=FALSE,
                         lambda=0,
                         xreg=fourier(ts_temp_USA2,
                                      K=c(2,12)))
ARIMA_Four_for <- forecast(ARIMA_Four,
                           xreg=fourier(ts_temp_USA2,
                                        K=c(2,12),
                                        h=365),
                           h=365)

#### NEURAL NETWORK 
NN_fit <- nnetar(ts_temp_USA2,p=1,P=1)

#5 year forecast
NN_for <- forecast(NN_fit, h=365)


### TBATS
TBATS_fit <- tbats(ts_temp_USA2)

TBATS_for <- forecast(TBATS_fit, h=365)

### STL + ETS
ETS_fit <- stlf(ts_temp_USA2, h=365)


###PLOT
autoplot(window(ts_temp_USA, start = c(2019,1))) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  autolayer(ARIMA_forecast, series="ARIMA",PI=FALSE)+
  autolayer(TBATS_for,series="TBATS",PI=FALSE)+
  autolayer(ETS_fit,series="STL + ETS",PI=FALSE)+
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE)+
  ylab("Temperature") 


a_score <- accuracy(ARIMA_forecast)
a2_score <- accuracy(ARIMA_Four_for)
NN_score <- accuracy(NN_for)
TBATS_score <- accuracy(TBATS_for)
ETS_score <- accuracy(ETS_fit)


scores <- as.data.frame(rbind(a_score, a2_score, NN_score, TBATS_score, ETS_score))
row.names(scores) <- c("ARIMA", "ARIMA FOURIER","NN", "TBATS", "STL + ETS")
scores

kbl(scores, 
      caption = "Forecast Accuracy for USA Temperature Data",
      digits = array(5,ncol(scores))) %>%
  kable_styling(full_width = FALSE, position = "center", html_font="Cambria")
```

##Discussion

The positive linear temperature trends for Brazil, United States, and Australia indicate the presence of climate change effects (Seasonal Mann-Kendall: p-values < 0.05). Temperature trends for the three countries follow a similar visual trend No one country looks to be increasing in temperature quicker than the other two. 

The ARIMA + Fourier and TBATS model are generally the best predictors for temperature changes over time as displayed in the error terms. However, looking at the forecasts, the STL + ETS looks to do best in all three countries. This is not reflected in the error terms. One explanation for this might be because the accuracy function only uses the training set to provide the accuracy scores. It cannot get the accuracy score for forecasted values that do not have observed values yet. For this reason, the test sets might provide a more accurate representation of which forecasting model performs the best. In the test accuracy function, we were able to provide the test set error along with the training set error to provide the most accurate error scores. 

Future steps might include looking into exogenous variables or events related to temperature such as precipitation, El Nino, and La Nina. Precipitation and temperature are closely linked and might provide a more accurate forecast model. 

Furthermore, it might be interesting to model precipitation or model more localized temperature trends in the future. Particularly, climate change has changed precipitation regimes, and would be vital to understand these trends in the context of climate risks and vulnerability. We looked at temperature trends over a national scale, however temperature trends may be masked at this granularity. Future forecasting studies may delve into localized temperature predictions.

##Limitations

Temperatures were generalized for the entire countries analyzed in this project. This approach can cause issues since regional climates in countries can contrast significantly. Countries were also selected bases on personal biases. Although daily data was available, only monthly temperatures were forecasted. Daily temperature would allow for a finer tuned model that would pick up weekly trends. Error terms were compared amongst forecasts, which only gave us a limited understanding of how good the forecast was. 
